---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# iod25

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
<!-- badges: end -->

The goal of iod25 is to provide R users with easier access to the English Indices of Deprivation 2025.

## Installation

``` r
# install.packages("pak")
pak::pak("douglasmunro/iod25")
```

## About the data

The Indices of Deprivation 2025 (IoD25) measure relative deprivation across small geographic areas in England.

IoD25 is formed from 7 domains which measure different aspects of deprivation, one combined Index of Multiple Deprivation (IMD) and 2 supplementary indices.

The iod25 package is made up of 3 core datasets:

- `domains` contains the 7 domain measures
- `imd` contains the Index of Multiple Deprivation
- `supplementary` contains the 2 supplementary indices

And two additional datasets:

- `subdomains` contains the underlying measures used to construct the domains
- `population` contains the population denominators used to construct the indices

## License

The package data are available under [CC BY 4.0](https://creativecommons.org/licenses/by/4.0/deed.en) license and are modified versions of Ministry of Housing, Communities &
Local Government (MHCLG) data.

The source data is licenced under [Open Government Licence v3.0. (OGLv3.0)](https://www.nationalarchives.gov.uk/doc/open-government-licence/version/3/).


## Disclaimer

This package is not associated with or endorsed by MHCLG.

## Citation

To cite the iod25 package, please use:

```{r}
citation("iod25")
```

To cite the source data, please use:

+ Ministry of Housing, Communities &
Local Government (MHCLG), Indices of Multiple Deprivation 2025, (2025). Available at: [https://open-data.communities.gov.uk/datasets/indices-of-deprivation-2025/](https://open-data.communities.gov.uk/datasets/indices-of-deprivation-2025/) (Accessed: 25 May 2025). Licensed under the [Open Government Licence v3.0](https://www.nationalarchives.gov.uk/doc/open-government-licence/version/3/).